 

### 2021-05-23 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Exposed-sun-leaves, though equipped with thermoregulatory traits are largely constrained by $gs$ than shade-leaves [@sanchesDifferentialLeafTraits2010; @hernandezSimilarTemperatureDependence2020]. Where evapotranspiration and $gs$ peaks during the day [@kosugiSeasonalFluctuationsTemperature2006; @sanchesDifferentialLeafTraits2010] and declines during midday and afternoon due to stomtal closure [@sanchesDifferentialLeafTraits2010; @zwienieckiHydraulicLimitationsImposed2004; @kochDiurnalPatternsLeaf1994b]. While specie-specific differences in $T_{opt}$ of $gs$ can influence their photosynthetic heat tolerance; exactly how, remains understudied [@perezPhotosyntheticHeatTolerances2020].

@@

$T_{leaf}$ beyond photosynthetic $T_{opt}$, declines photosynthesis due it's peaked response to temperature. Photosynthesis is commonly found to be maximized at the prevalent ambient growing season conditions (Tan et al. 2017; @slotSituTemperatureResponse2017). Beyond the optimum, photosynthesis decreases as a result of stomatal closure and increasing $T_{leaf}$ (e.g. @slotSituTemperatureResponse2017; Smith et al. 2020 [temperate/boreal refs needed) and eventually due to biochemical constraints (@kumarathungeAcclimationAdaptationComponents2019a; e.g. Sage & Kubien 2007; Varhammer et al. 2015 ). Leaf heat stress at $40-60{^\circ}C$ can cause irreversible damage to photosystem [II] leading to leaf necrosis and eventually leaf death [Baker 2008; @feeley_thermal_2020].

>>>>

Exposed-sun-leaves, though equipped with thermoregulatory traits are largely constrained by $gs$ than shade-leaves [@sanchesDifferentialLeafTraits2010; @hernandezSimilarTemperatureDependence2020]. Where evapotranspiration and $gs$ peaks during the day [@kosugiSeasonalFluctuationsTemperature2006; @sanchesDifferentialLeafTraits2010] and declines during midday and afternoon due to stomtal closure [@sanchesDifferentialLeafTraits2010; @zwienieckiHydraulicLimitationsImposed2004; @kochDiurnalPatternsLeaf1994b]. While specie-specific differences in $T_{opt}$ of $gs$ can influence their photosynthetic heat tolerance indirectly; exactly how, remains understudied [@perezPhotosyntheticHeatTolerances2020].

@@

Temperature can affect photosynthesis via direct and indirect pathways. Direct, by altering photosynthetic enzyme activity and electron transport chain, and indirectly through increase in $T_{leaf}$ from $gs$ closure that is driven by biophysical factors [@lloydEffectsRisingTemperatures2008]. $T_{leaf}$ beyond photosynthetic $T_{opt}$, declines photosynthesis due it's peaked response to temperature. Photosynthesis is commonly found to be maximized at the prevalent ambient growing season conditions (Tan et al. 2017; @slotSituTemperatureResponse2017). Beyond the optimum, photosynthesis decreases as a result of stomatal closure and increasing $T_{leaf}$ (e.g. @slotSituTemperatureResponse2017; Smith et al. 2020 [temperate/boreal refs needed) and eventually due to biochemical constraints (@kumarathungeAcclimationAdaptationComponents2019a; e.g. Sage & Kubien 2007; Varhammer et al. 2015 ). Leaf heat stress at $40-60{^\circ}C$ can cause irreversible damage to photosystem [II] leading to leaf necrosis and eventually leaf death [Baker 2008; @feeley_thermal_2020].
